Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy

### 2020-07-17 chapters/ch_results.tex
Almost final draft

\chapter{Results}
%\section{Comparison to Theory}
%Add figure of Pion spectra for different theory 
%Reference paper for multiple theories

@@

\begin{table*}\centering
\ra{1.3}
\begin{tabular}{@{}cccc@{}}\toprule
System & $\pi^-$ & $\pi^+$ & $Y(\pi^-)/Y(\pi^+)$  \\
\midrule
$\tin{132}{124}$ & 0.717(24)(4) & 0.148(5)(2) & 4.84(10)(6)  \\
$\tin{108}{112}$ & 0.399(14)(3) & 0.200(8)(2) & 1.99(4)(3)  \\
%$\tin{132}{124}$ & \numerr{0.717}{0.024}{0.004} & \numerr{0.148}{0.005}{0.002} & \numerr{4.84}{0.10}{0.06}  \\
%$\tin{108}{112}$ & \numerr{0.399}{0.014}{0.003} & \numerr{0.200}{0.008}{0.002} & \numerr{1.99}{0.04}{0.03}  \\
\bottomrule
\end{tabular}
\caption{Total pion yield.}
\label{tb:pionyield}
\end{table*}

@@

The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher, beam energies. In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$. Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Fig.~\ref{fig:totalpiRatio}. Here the dependence is fitted with a 2-nd order polynomial fit. To compare the pion ratio in the $\tin{132}{124}$ data with that of the lower N/Z in the FOPI experiment, we scaled by the pion ratio between $N/Z=1.56$ and 1.493 as given from the fitted polynomial line. Figure~\ref{fig:fopiPionRatio} shows the scaled pion ratio as compared with the FOPI pion ratio \cite{fopi}. The fitted function has the functional form of $p_1(E - p_2)^{-2}$ where $p_1$ and $p_2$ are free parameters, and only is meant to guide the eye. It is also worth mentioning that the pion ratio observed in the FOPI \SI{400}{\MeVA} setting was already considerably higher than what is expected from the $(N/Z)^2$ delta resonance model \cite{baoan_piprod1,baoan_piprod2}. It it worth mentioning that the two experiments differ considerably in their efficiencies as was mentioned before. The FOPI results heavily relied on extrapolations to lower pion energies which the \spirit TPC did not rely on \cite{fopi}. Without a systematic error calculation or discussion with the FOPI collaboration no further detail can be given except that the \spirit TPC data seems to be reasonable expectation and in general agreement with the FOPI data set. Scaling the other systems in the \spirit TPC data leads to almost the exact same scaled value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$ and would be redundant.

@@

\begin{comment}
\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{pionSpectra.png}
\caption{Pion spectra. }
\label{fig:pionspectra}
\end{figure}
%need to add error analysis
\end{comment}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{pionSpectra_sn108_sum.png}
\caption{Pion spectra $\tin{108}{112}$ system. }
\label{fig:pionspectraSn108}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{pionSpectra_sn132_sum.png}
\caption{Pion spectra $\tin{132}{124}$ system. }
\label{fig:pionspectraSn132}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{singleRatio108_log.png}
\caption{Single ratio $\tin{108}{112}$ system.}
\label{fig:SRsn108}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{singleRatio132_sum.png}
\caption{Single ratio $\tin{132}{124}$ system.}
\label{fig:SRsn132}
\end{figure}

>>>>

\chapter{Results}
\label{chap:results}

@@

\begin{table*}\centering
\ra{1.3}
\begin{tabular}{@{}cccc@{}}\toprule
System & $\pi^-$ & $\pi^+$ & $Y(\pi^-)/Y(\pi^+)$  \\
\midrule
$\tin{132}{124}$ & 0.717(24)(4) & 0.148(5)(2) & 4.84(10)(6)  \\
$\tin{108}{112}$ & 0.399(14)(3) & 0.200(8)(2) & 1.99(4)(3)  \\
%$\tin{132}{124}$ & \numerr{0.717}{0.024}{0.004} & \numerr{0.148}{0.005}{0.002} & \numerr{4.84}{0.10}{0.06}  \\
%$\tin{108}{112}$ & \numerr{0.399}{0.014}{0.003} & \numerr{0.200}{0.008}{0.002} & \numerr{1.99}{0.04}{0.03}  \\
\bottomrule
\end{tabular}
\caption{Measured total pion yield in the $\tin{132}{124}$ and $\tin{108}{112}$ systems.}
\label{tb:pionyield}
\end{table*}

@@

The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher, beam energies. In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$. Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Figure~\ref{fig:totalpiRatio}. Here the dependence is fitted with a 2-nd order polynomial fit. To compare the pion ratio in the $\tin{132}{124}$ data with that of the lower N/Z in the FOPI experiment, we scaled by the pion ratio between $N/Z=1.56$ and 1.493 as given from the fitted polynomial line. Figure~\ref{fig:fopiPionRatio} shows the scaled pion ratio as compared with the FOPI pion ratio \cite{fopi}. The fitted function has the functional form of $p_1(E - p_2)^{-2}$ where $p_1$ and $p_2$ are free parameters, and only is meant to guide the eye. It is also worth mentioning that the pion ratio observed in the FOPI \SI{400}{\MeVA} setting was already considerably higher than what is expected from the $(N/Z)^2$ delta resonance model \cite{baoan_piprod1,baoan_piprod2}. It it worth mentioning that the two experiments differ considerably in their efficiencies as was mentioned before. The FOPI results heavily relied on extrapolations to lower pion energies which the \spirit TPC did not rely on \cite{fopi}. Without a systematic error calculation or discussion with the FOPI collaboration no further detail can be given except that the \spirit TPC data seems to be reasonable expectation and in general agreement with the FOPI data set. Scaling the other systems in the \spirit TPC data leads to almost the exact same scaled value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$ and would be redundant.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{pionSpectra_sn108_sum.png}
\caption{Pion spectra for the $\tin{108}{112}$ system and comparisons to 7 theoretical models. }
\label{fig:pionspectraSn108}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{pionSpectra_sn132_sum.png}
\caption{Pion spectra for the $\tin{132}{124}$ system and comparisons to 7 theoretical models. }
\label{fig:pionspectraSn132}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{singleRatio108_log.png}
\caption{Pion spectral ratio $\tin{108}{112}$ system and comparisons to 7 theoretical models.}
\label{fig:SRsn108}
\end{figure}

\begin{figure}[!htb]
\centering
\includegraphics[width=\textwidth]{singleRatio132_sum.png}
\caption{Pion spectral ratio $\tin{132}{124}$ system and comparisons to 7 theoretical models.}
\label{fig:SRsn132}
\end{figure}

### 2020-07-17 chapters/ch_summary.tex
Almost final draft

\chapter{Summary}
The need for precise experimental constraints on the symmetry energy at high densities is paramount to understanding exotic, dense matter, such as neutron stars. Direct observation of these objects provides information about the total equation of state, which is primarily sensitive to the pressure in neutron matter. Only by laboratory measurements can one control the asymmetry of matter and isolate its dependence on the isospin asymmetry $\delta = \frac{\rho_n - \rho_p}{\rho}$. This information is critical to understanding the symmetry energy and associated isovector mean field potentials. These mean field potential contribute strongly to the chemical potentials, which control the phase transitions and the internal structure within stars. Heavy ion collisions provide the only laboratory environment to study these issues at the suprasaturation densities relevant to the neutron star interior.

>>>>

\chapter{Summary}
\label{chap:summary}

The need for precise experimental constraints on the symmetry energy at high densities is paramount to understanding exotic, dense matter, such as neutron stars. Direct observation of these objects provides information about the total equation of state, which is primarily sensitive to the pressure in neutron matter. Only by laboratory measurements can one control the asymmetry of matter and isolate its dependence on the isospin asymmetry $\delta = \frac{\rho_n - \rho_p}{\rho}$. This information is critical to understanding the symmetry energy and associated isovector mean field potentials. These mean field potential contribute strongly to the chemical potentials, which control the phase transitions and the internal structure within stars. Heavy ion collisions provide the only laboratory environment to study these issues at the suprasaturation densities relevant to the neutron star interior.
